Limb-darkening and exoplanets II: Choosing the Best Law for Optimal Retrieval of Transit Parameters

### 2016-01-18 which_law_2_use.tex
Added extra discussion to explain MSE results.

In the past decade, the study of transiting exoplanets has evolved from discovery to very precise characterization of these systems, 
thanks to the exquisite precision allowed mainly by space-based observatories such as the {\em Hubble Space Telescope} ({\em HST}) and the {\em Kepler} mission \citep{koch2010}.
This has allowed the study of precise mass-radius diagrams \citep[see, e.g.,][and references therein]{wm2014}, and in turn  allow the possibility of not only obtaining a precise determination of the internal composition of small, rocky exoplanets based on mass-radius
measurements \citep{dorn2015} or of their atmospheric fractions from radius measurements alone \citep{WL2015}, but also of the determination of
derived parameters of the systems that even allow for the estimation of stellar parameters directly from transit lightcurves through techniques such as asterodensity
profiling \citep{sm2003,kipping2014AP}, or of the detection of atmospheric features in exoplanet atmospheres through the technique of transmission spectroscopy. 
All these studies and techniques rely on the retrieval of precise and accurate transit parameters from transit lightcurves.

@@

From the results on the past sub-section, one can see that for a given set of stellar and transit parameters, a law can be chosen in order to minimize the 
biases in the retrieved transit parameters from the transit light curve. However, an interesting question is whether the precision of a given lightcurve can 
actually provide one with enough precision in order to use a more complex parametrization of the limb-darkening effect and, depending on the answer, what is 
the cost. Fitting a complex law to a noisy light curve will definitely lower the precision on the retrieved transit parameters due to the high flexibility of the 
model (i.e., the large parameter space that it can explore), while fitting a law that is too simple for a very precise lightcurve will most definitely lead to larger 
biases, due to the strictness of the chosen model. This problem is known as the bias-variance trade-off, i.e., the problem of choosing the right amount of bias 
and variance for a given dataset. Probabilistically, for a given estimator $\hat{\theta}$ of the parameter $\theta$, we will be interested in its variance (whose 
square-root defines the precision), $\textnormal{Var}[\hat{\theta}]$, and its bias, defined as $\mathbb{E}[\hat{\theta}]-\theta$, where 
$\mathbb{E}[\cdot]$ defines the expectation value and $\textnormal{Var}[\cdot]$ is the variance operator. The bias, on one hand, defines how far on average is 
the estimator from the parameters' true value, while the variance defines the spread around the expected value of the estimator (see Figure \ref{bias-variance-plot}). 
In this sense, one would ideally want a low-bias and high-precision (i.e. low variance) estimator, which is both close to the real value of the parameter on average and, 
at the same time, has a low spread around such value. The problem is that in real applications one has the possibility to choose from various estimators with different 
ammounts of bias and variance, and thus choosing among them is, at first, not trivial. This is when the mean-squared error metric comes in handy. The mean-squared 
error (MSE) of an estimator $\hat{\theta}$ of the parameter $\theta$ is defined as
\begin{eqnarray*}
\textnormal{MSE} &=& \mathbb{E}[(\theta - \hat{\theta})^2].
\end{eqnarray*}
From the definition of the MSE, it can be seen that it is a measure of the distance from the estimator to the parameters' ``true" value. As such, this provides an 
excellent metric among a given suite of estimators for a given parameter: for a given set of estimators, the one that gives the minimum mean-squared error 
provides, on average, the one that is closer to the underlying ``true" value on a mean-square sense. Furthermore, it is easy to shown that the MSE can be written 
as
\begin{eqnarray*}
\textnormal{MSE} &=& \mathbb{E}[(\theta - \hat{\theta})^2], \\
&=& (\mathbb{E}[\hat{\theta}]-\theta)^2 + \textnormal{Var}[\hat{\theta}],\\
&=& \textnormal{Bias}^2 + \textnormal{Variance}.
\end{eqnarray*}
This implies that given the bias and the variance of an estimator, the MSE can be easily calculated. In our specific application, in the past sub-section we have already 
shown and studied the biases that each limb-darkening law (i.e., different estimators) imply on the retrieved transit parameters through simulation. We now estimate 
the variances implied by those limb-darkening laws.

@@

Given that the study of the mean-squared error is complex from a mathematical point of view in this case, we study it through simulations similar to the ones done in the past sub-section, 
but this time also adding different amounts of noise to the lightcurves in order to estimate the variances of the estimators. In this sense, we take a Monte Carlo approach in order to study 
the MSE. As a case study, we take that of a typical Hot-Jupiter ($p=0.1$, $a_R = 10$) around a Solar-type star ($T_\textnormal{eff}= 5500$ K) on a low impact-parameter orbit ($b=0.3$), 
and we focus specifically on the retrieval of the planet-to-star radius ratio, which is one of the most interesting parameters retrieved from transit lightcurves. We simulate a suite of 
lightcurves with noise levels ranging from $10$ ppm to $3000$ ppm with different amount of in-transit points; here, however, we show our results for $N=100$ and $N=1000$ in-transit 
points for clarity, which span a range of possible scenarios for both space based (e.g., \textit{Kepler}) and ground-based observations. For each case, $300$ lightcurves are generated, 
and all of them are fitted in the same way as it was done for their noiseless counterparts in the past sub-section. From the fit to each of the $300$ lightcurves on each case, 
we take all the estimated planet-to-star radius ratios and estimate the variance of the sample. The results of our simulations are shown on Figure \ref{biasvariance}, where we plot the 
MSE as a function of the noise in the simulated lightcurves.

In our results, it can be seen that the optimal law in terms of the MSE at each noise level can be selected easily. For example, in our simulation, for $N=100$ in-transit points, 
the logarithmic law is the best for the retrieval of the planet-to-star radius ratio in a MSE sense from lightcurves with noise levels of $\sim3000$ ppm up to $\sim 1000$ ppm. 
At this last noise level, it can be seen that the square-root law crosses the line of the logarithmic law, and this law attains the lowest MSE until $\sim 200$ ppm, where the line 
quadratic law starts to get the lowest MSE. The quadratic law is thus the best law to use from $\sim 200$ ppm until lightcurves with precisions of $\sim25$ ppm, precision at 
which the three-parameter law takes over and attains the lowest MSE. Note that the number of in-transit points changes where the transition among different limb-darkening 
laws occur. For example, the transition between the quadratic and three-parameter laws occurs with lower precision lightcurves for the case of $1000$ in-transit points 
(at $\sim 75$ ppm).

From our case study, it can be seen that the selection of the optimal limb-darkening law in a MSE sense can be estimated by simple simulation; however, we note that different 
transit geometries will give rise to different optimal limb-darkening laws at a given lightcurve precision and, thus, these simulations must be performed on a case-by-case basis. 
We provide code at Github with which this can be easily done for a given set of transit and stellar parameters, along with a given lightcurve noise level and number of in-transit points.

@@

\begin{thebibliography}{99}
\bibitem[\protect\citeauthoryear{Claret}{2000}]{claret2000} Claret, A., 2000, A\&A, 363, 1081.
\bibitem[\protect\citeauthoryear{Claret \& Hauschildt}{2003}]{claret2003} Claret, A. \& Hauschildt, P. H., 2003, 
A\&A, 412, 241.
\bibitem[\protect\citeauthoryear{D\'iaz-Cordov\'ez \& Gim\'enez}{1992}]{diazgimenez1992} D\'iaz-Cordov\'ez, J.,
\& Gim\'enez, A., 1992, A\&A, 259, 227.
\bibitem[\protect\citeauthoryear{Dorn et al.}{2015}]{dorn2015} Dorn, C., Khan, A., Heng, K., Alibert, Y., Connolly, J.,
Benz, W. \& Tackley, P., 2015, A\&A, 577, 83.
\bibitem[\protect\citeauthoryear{Kipping}{2013}]{kipping2013} Kipping, D. M., 2013, MNRAS, 435, 2152.
\bibitem[\protect\citeauthoryear{Kipping}{2014}]{kipping2014AP} Kipping, D. M., 2014, MNRAS, 440, 2164.
\bibitem[\protect\citeauthoryear{Kipping}{2015}]{kipping2015} Kipping, D. M., 2015, arXiv:1509.03483.
\bibitem[\protect\citeauthoryear{Klinglesmith \& Sobieski}{1970}]{klinglesmith1970} Klinglesmith, D. A. \&
Sobieski, S., 1970, AJ, 75, 175.
\bibitem[\protect\citeauthoryear{Espinoza \& Jord\'an}{2015}]{ej2015} Espinoza, N. \& Jord\'an, A., 2015, 
MNRAS, 450, 1879.
\bibitem[\protect\citeauthoryear{Howarth}{2011}]{howarth2011} Howarth, I., 2011, 
MNRAS, 413, 1515.
\bibitem[\protect\citeauthoryear{Klinglesmith \& Sobieski}{1970}]{klinglesmith1970} Klinglesmith, D. A. \&
Sobieski, S., 1970, AJ, 75, 175.
\bibitem[\protect\citeauthoryear{Knutson et al.}{2007}]{knutson2007} Knutson, H., Charbonneau, D., Noyes, R., et al., 2007, ApJ, 655, 564.

@@

\bibitem[\protect\citeauthoryear{Seager \& Mall\'en-Ornelas}{2003}]{sm2003} Seager, S. \& Mall\'en-Ornelas, G., 2003, ApJ, 585, 1038.
\bibitem[\protect\citeauthoryear{Sing et al.}{2009}]{sing2009} Sing, D. K., D\'esert, J.-M.,
Lecavelier Des Etangs, A., Ballester, G. E., Vidal-Madjar, A., Parmentier, V., Hebrard, G. \& Henry, G. W.,
2009, A\&A, 505, 891.
\bibitem[\protect\citeauthoryear{Weiss \& Marcy}{2014}]{wm2014} Weiss, L. \& Marcy, G., 2014, ApJ, 783L, 6.
\bibitem[\protect\citeauthoryear{Wolfgang \& Lopez}{2015}]{WL2015} Wolfgang, A. \& Lopez, E., 2015, ApJ, 806, 183.
\end{thebibliography}

>>>>

In the past decade, the study of transiting exoplanets has evolved from discovery to very precise characterization of these systems, 
thanks to the exquisite precision allowed mainly by space-based observatories such as the {\em Hubble Space Telescope} ({\em HST}) and the {\em Kepler} mission \citep{koch2010}.
This has allowed the study of precise mass-radius diagrams \citep[see, e.g.,][]{wm2014,dressing2015,WL2015b}, and in turn  allow the possibility of not only obtaining a precise determination of the internal composition of small, rocky exoplanets based on mass-radius
measurements \citep{dorn2015} or of their atmospheric fractions from radius measurements alone \citep{WL2015}, but also of the determination of
derived parameters of the systems that even allow for the estimation of stellar parameters directly from transit lightcurves through techniques such as asterodensity
profiling \citep{sm2003,kipping2014AP}, or of the detection of atmospheric features in exoplanet atmospheres through the technique of transmission spectroscopy. 
All these studies and techniques rely on the retrieval of precise and accurate transit parameters from transit lightcurves.

@@

From the results on the past sub-section, one can see that for a given set of stellar and transit parameters, a law can be chosen in order to minimize the 
biases in the retrieved transit parameters from the transit light curve. However, an interesting question is whether the precision of a given lightcurve can 
actually provide one with enough precision in order to use a more complex parametrization of the limb-darkening effect and, depending on the answer, what is 
the cost. Fitting a complex law to a noisy light curve will definitely lower the precision on the retrieved transit parameters due to the high flexibility of the 
model (i.e., the large parameter space that it can explore), while fitting a law that is too simple for a very precise lightcurve will most definitely lead to larger 
biases, due to the strictness of the chosen model. This problem is known as the bias-variance trade-off, i.e., the problem of choosing the right amount of bias 
and variance for a given dataset. Probabilistically, for a given estimator $\hat{\theta}$ of the parameter $\theta$, we will be interested in its variance (whose 
square-root defines the precision), $\textnormal{Var}[\hat{\theta}]$, and its bias, defined as $\mathbb{E}[\hat{\theta}]-\theta$, where 
$\mathbb{E}[\cdot]$ defines the expectation value and $\textnormal{Var}[\cdot]$ is the variance operator. The bias, on one hand, defines how far on average is 
the estimator from the parameters' true value, while the variance defines the spread around the expected value of the estimator (see Figure \ref{bias-variance-plot}). 
In this sense, one would ideally want a low-bias and high-precision (i.e. low variance) estimator, which is both close to the real value of the parameter on average and, 
at the same time, has a low spread around such value. The problem is that in real applications one has the possibility to choose from various estimators with different 
ammounts of bias and variance, and thus choosing among them is, at first, not trivial. This is when the mean-squared error metric comes in handy. The mean-squared 
error (MSE) of an estimator $\hat{\theta}$ of the parameter $\theta$ is defined as
\begin{eqnarray*}
\textnormal{MSE} &=& \mathbb{E}[(\theta - \hat{\theta})^2].
\end{eqnarray*}
From the definition of the MSE, it can be seen that it is a measure of the square of the distance from the estimator to the parameters' ``true" value. As such, this provides an 
excellent metric among a given suite of estimators for a given parameter: for a given set of estimators, the one that gives the minimum mean-squared error 
provides, on average, the one that is closer to the underlying ``true" value on a mean-square sense. Furthermore, it is easy to show that the MSE can be written 
as
\begin{eqnarray*}
\textnormal{MSE} &=& \mathbb{E}[(\theta - \hat{\theta})^2], \\
&=& (\mathbb{E}[\hat{\theta}]-\theta)^2 + \textnormal{Var}[\hat{\theta}],\\
&=& \textnormal{Bias}^2 + \textnormal{Variance}.
\end{eqnarray*}
This implies that given the bias and the variance of an estimator, the MSE can be easily calculated. In our specific application, in the past sub-section we have already 
shown and studied the biases that each limb-darkening law (i.e., different estimators) imply on the retrieved transit parameters through simulation. In this sense, we have 
calculated the floor of the MSE for each law, and it only remains to study the variances implied by those limb-darkening laws, which is a function of the noise of the lightcurve 
used to retrieve the transit parameters.

@@

Given that the study of the mean-squared error is complex from a mathematical point of view in this case, we study it through simulations similar to the ones done in the past sub-section, 
but this time also adding different amounts of noise to the lightcurves in order to estimate the variances of the estimators. In this sense, we take a Monte Carlo approach in order to study 
the MSE. As a case study, we take that of a typical Hot-Jupiter ($p=0.1$, $a_R = 10$) around a Solar-type star ($T_\textnormal{eff}= 5500$ K) on a low impact-parameter orbit ($b=0.3$), 
and we focus specifically on the retrieval of the planet-to-star radius ratio, which is one of the most interesting parameters retrieved from transit lightcurves. We simulate a suite of 
lightcurves with noise levels ranging from $10$ ppm to $3000$ ppm with different amount of in-transit points; here, however, we show our results for $N=100$ and $N=1000$ in-transit 
points for clarity, which span a range of possible scenarios for both space based (e.g., \textit{Kepler}) and ground-based observations. For each case, $300$ lightcurves are generated, 
and all of them are fitted in the same way as it was done for their noiseless counterparts in the past sub-section. From the fit to each of the $300$ lightcurves on each case, 
we take all the estimated planet-to-star radius ratios and estimate the variance of the sample. The results of our simulations are shown on Figure \ref{biasvariance}, where we plot the 
MSE as a function of the noise in the simulated lightcurves for each limb-darkening law used to retrieve the parameter.

In our results, it is evident how the MSE evolves from being variance-dominated for low-precision lightcurves (rightmost part of the plot) to bias-dominated for high-precision 
lightcurves (leftmost part of the plot). This evolution is most evident for the linear law: for $N=100$ in-transit points, for example, the linear law attains the bias floor around 
$\sim100$ ppm. From here and for more precise lightcurves, the MSE of this law attains the constant value $\sim 10^{-6}$, which corresponds to the bias floor of 
$10^{-3}$, which for $p=0.1$ corresponds to the $1\%$ bias observed on our simulations on the past sub-section (Figure \ref{small_b_sims}). 
Note that from this plot, the optimal law in terms of the MSE at each noise level can be selected easily. For example, in our simulation, for $N=100$ in-transit points, 
the logarithmic law is the best for the retrieval of the planet-to-star radius ratio in a MSE sense from lightcurves with noise levels of $\sim3000$ ppm up to $\sim 1000$ ppm. 
At this last noise level, it can be seen that the square-root law crosses the line of the logarithmic law, and this law attains the lowest MSE until $\sim 200$ ppm, where the 
quadratic law starts to get the lowest MSE. The quadratic law is thus the best law to use from $\sim 200$ ppm until lightcurves with precisions of $\sim25$ ppm, precision at 
which the three-parameter law takes over and attains the lowest MSE. Note that the number of in-transit points changes where the transition among different limb-darkening 
laws occur due to the fact that the bias floor is attained faster due to the increased number of points. For example, the transition between the quadratic and three-parameter 
laws occurs with lower precision lightcurves for the case of $1000$ in-transit points (at $\sim 75$ ppm).

From our case study, it can be seen that the selection of the optimal limb-darkening law in a MSE sense can be estimated by simple simulation: simply calculate the MSE obtained 
by all the laws and select the one that gives the minimum MSE. However, we note that different transit geometries will give rise to different optimal limb-darkening laws at a given 
lightcurve precision and, thus, these simulations must be performed on a case-by-case basis. Of course, for this case-by-case study a set of transit parameters are needed as 
inputs, which might be strange as this is what we are trying to extract from a transit lightcurve. However, as we learned from our results on this and the past sub-sections, it is important 
to remember that the functional relation of the transit parameters and the biases/variances is smooth and slow; (i.e., small changes on the retrieved transit parameters give rise to 
similar biases and variances). As such, one might first one to fit with one favourites' law, obtain a set of transit parameters and use these as inputs in order to calculate the MSE of 
different limb-darkening laws in order to retrieve the final transit parameters from the law that gives the lowest MSE. We provide code at Github with which this can be easily done 
for a given set of transit and stellar parameters, along with a given lightcurve noise level and number of in-transit points.

@@

\begin{thebibliography}{99}
\bibitem[\protect\citeauthoryear{Claret}{2000}]{claret2000} Claret, A., 2000, A\&A, 363, 1081.
\bibitem[\protect\citeauthoryear{Claret \& Hauschildt}{2003}]{claret2003} Claret, A. \& Hauschildt, P. H., 2003, 
A\&A, 412, 241.
\bibitem[\protect\citeauthoryear{D\'iaz-Cordov\'ez \& Gim\'enez}{1992}]{diazgimenez1992} D\'iaz-Cordov\'ez, J.,
\& Gim\'enez, A., 1992, A\&A, 259, 227.
\bibitem[\protect\citeauthoryear{Dorn et al.}{2015}]{dorn2015} Dorn, C., Khan, A., Heng, K., Alibert, Y., Connolly, J.,
Benz, W. \& Tackley, P., 2015, A\&A, 577, 83.
\bibitem[\protect\citeauthoryear{Dressing et al.}{2015}]{dressing2015} Dressing, C. D., Charbonneau, D., Dumusque, X., et al., 2015, ApJ, 800, 135.
\bibitem[\protect\citeauthoryear{Kipping}{2013}]{kipping2013} Kipping, D. M., 2013, MNRAS, 435, 2152.
\bibitem[\protect\citeauthoryear{Kipping}{2014}]{kipping2014AP} Kipping, D. M., 2014, MNRAS, 440, 2164.
\bibitem[\protect\citeauthoryear{Kipping}{2015}]{kipping2015} Kipping, D. M., 2015, arXiv:1509.03483.
\bibitem[\protect\citeauthoryear{Klinglesmith \& Sobieski}{1970}]{klinglesmith1970} Klinglesmith, D. A. \&
Sobieski, S., 1970, AJ, 75, 175.
\bibitem[\protect\citeauthoryear{Espinoza \& Jord\'an}{2015}]{ej2015} Espinoza, N. \& Jord\'an, A., 2015, 
MNRAS, 450, 1879.
\bibitem[\protect\citeauthoryear{Howarth}{2011}]{howarth2011} Howarth, I., 2011, 
MNRAS, 413, 1515.
\bibitem[\protect\citeauthoryear{Klinglesmith \& Sobieski}{1970}]{klinglesmith1970} Klinglesmith, D. A. \&
Sobieski, S., 1970, AJ, 75, 175.
\bibitem[\protect\citeauthoryear{Knutson et al.}{2007}]{knutson2007} Knutson, H., Charbonneau, D., Noyes, R., et al., 2007, ApJ, 655, 564.

@@

\bibitem[\protect\citeauthoryear{Seager \& Mall\'en-Ornelas}{2003}]{sm2003} Seager, S. \& Mall\'en-Ornelas, G., 2003, ApJ, 585, 1038.
\bibitem[\protect\citeauthoryear{Sing et al.}{2009}]{sing2009} Sing, D. K., D\'esert, J.-M.,
Lecavelier Des Etangs, A., Ballester, G. E., Vidal-Madjar, A., Parmentier, V., Hebrard, G. \& Henry, G. W.,
2009, A\&A, 505, 891.
\bibitem[\protect\citeauthoryear{Weiss \& Marcy}{2014}]{wm2014} Weiss, L. \& Marcy, G., 2014, ApJ, 783L, 6.
\bibitem[\protect\citeauthoryear{Wolfgang \& Lopez}{2015}]{WL2015} Wolfgang, A. \& Lopez, E., 2015, ApJ, 806, 183.
\bibitem[\protect\citeauthoryear{Wolfgang, Rogers \& Ford}{2015}]{WL2015b} Wolfgang, A., Rogers, L. \& Ford, E., 2015, eprint arXiv:1504.07557.
\end{thebibliography}
